---
title: "Visualizing National park_visits"
author: "Freaky Fritters"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Our National park_visits

America has 61 national park_visits. They have grown in size and popularity over the years.

```{r loading-libraries, message=F}
library(devtools)
library(tidyverse)
library(gganimate)
library(gifski)
library(png)
library(here)
library(lemon)
library(RColorBrewer)
library(scales)
library(janitor)
library(ggrepel)
library(ggvoronoi)
# library(ggtext)
library(rvest)
library(fuzzyjoin)
library(jkmisc)
library(ragg)
library(plotly)
```

## Code to clean and tidy the original data
This is a demostration of using the `janitor` package

```{r cleaning-data, include=F}
df_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/All%20National%20Parks%20Visitation%201904-2016.csv")


df <- df_raw %>% 
  janitor::clean_names() %>%
  mutate(date = lubridate::mdy_hms(year)) %>% 
  select(date, gnis_id, geometry:year_raw)

df %>% 
  write_csv(here::here("data/national_parks.csv"))


# Get pop data

url <- "https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_historical_population"

raw_html <- url %>% 
  read_html() %>% 
  html_table()

pop_df <- raw_html %>% 
  chuck(5) %>% 
  gather(key = "state", value = "pop", AL:DC) %>% 
  rename("year" = 1) %>% 
  mutate(pop = str_remove_all(pop, ","),
         pop = as.double(pop))

pop_df %>% 
  write_csv(here::here("data", "state_pop.csv"))

# Get gas prices

url2 <- "https://www.energy.gov/eere/vehicles/fact-915-march-7-2016-average-historical-annual-gasoline-pump-price-1929-2015"

raw_gas <- url2 %>% 
  read_html() %>% 
  html_table()

gas <- raw_gas %>% 
  chuck(1) %>% 
  set_names(nm = c("year", "gas_current", "gas_constant")) %>%   
  as_tibble() %>% 
  filter(!str_detect(year, "Source")) %>% 
  mutate(year = as.double(year),
         gas_current = as.double(gas_current),
         gas_constant = as.double(gas_constant))

##Added by Team FF
gas %>%
  write_csv(here::here("data", "national_parks.csv"))

```


## These visuals can help see the change in popularity

We will now import the datasets as prepared and cleaned by the tidytuesday folks

```{r import-dataset, message=FALSE}
park_visits <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")
state_pop <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/state_pop.csv")
gas_price <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/gas_price.csv")
park_locations <- read_csv("data/NationalParksLatLong.csv")
```
# Data Dictionary

## `national_parks.csv`

|variable          |class     |description |
|:-----------------|:---------|:-----------|
|year_raw          |integer | Year of record |
|gnis_id           |character | ID for shapefile and long-lat lookup |
|geometry          |character | Geometric shape for shapefile |
|metadata          |character | URL to metadata about the park |
|number_of_records |double    | Number of records |
|parkname          |character | Full park name |
|region            |character | US Region where park is located |
|state             |character | State abbreviation |
|unit_code         |character | Park code abbreviation |
|unit_name         |character | Park Unit name |
|unit_type         |character | Park unit type |
|visitors          |double    | Number of visitors |

## `state_pop.csv`

|variable |class     |description |
|:--------|:---------|:-----------|
|year     |integer   | Jan 1st of year |
|state    |character | State abbreviation |
|pop      |double    | Population |

## `gas_price.csv`

|variable     |class  |description |
|:------------|:------|:-----------|
|year         |double | Year (Jan 1st) |
|gas_current  |double | Gas price in that year (dollars/gallon) |
|gas_constant |double | Gas price (constant 2015 dollars/gallon) |


Source: https://github.com/spren9er/tidytuesday/blob/master/tidytuesday_201938_national_park_visits.r
Improvments seen below are discussed here: https://evancanfield.netlify.com/posts/2019-09-16-national-parks-tidytuesday/tidy-tuesday-38/

```{r most-popular-national-park-visits}

park_visits <- park_visits %>%
  mutate(
    parkname =
      if_else(
        is.na(parkname),
        str_trim(str_remove(unit_name, 'National Park')),
        parkname
      )
  )

highlight_parks <- c(
  'GREAT SMOKY MOUNTAINS',
  'GRAND CANYON',
  'ROCKY MOUNTAIN',
  'YOSEMITE',
  'YELLOWSTONE',
  'ZION',
  'ACADIA',
  'DENALI',
  'HOT SPRINGS',
  'CARLSBAD CAVERNS',
  'GREAT BASIN'
)

highlight_colors <- c(
  '#223e15',
  '#176785',
  '#499989',
  '#5fa73f',
  '#ff8706',
  '#ff534e',
  '#f5b901',
  '#9a91fa',
  '#c988d2',
  '#6da5c2',
  '#fe43bc'
)

ranking_parks <- park_visits %>%
  filter(
    year != 'Total',
    unit_type == 'National Park',
    !is.na(parkname),
    unit_name != 'Denali National Preserve'
  ) %>%
  mutate(
    year = as.integer(year),
    parkname = str_to_upper(parkname)
  ) %>%
  filter(year < 2016) %>%
  group_by(year) %>%
  arrange(year, desc(visitors)) %>%
  mutate(rank = row_number()) %>%
  ungroup()

top_parks <- ranking_parks %>%
  filter(parkname %in% highlight_parks) %>%
  mutate(parkname = fct_relevel(str_to_upper(parkname), highlight_parks)) %>%
  arrange(year, desc(parkname))

other_parks <- ranking_parks %>%
  filter(!parkname %in% highlight_parks)

animation <- top_parks %>%
  ggplot(aes(x = year, y = rank, group = parkname, color = parkname)) +
  geom_line(
    data = other_parks, size = 0.5, show.legend = FALSE, color = '#dadada'
  ) +
  geom_line(show.legend = FALSE, size = 0.8) +
  geom_text(
    aes(x = year + 0.8, label = parkname),
    size = 4.5, show.legend = FALSE, hjust = 0, fontface = 'bold'
  ) +
  scale_x_continuous(breaks = c(1925, 1950, 1975, 2000)) +
  scale_y_continuous(
    breaks = c(1, 25, 50), labels = c('1ˢᵗ', '25ᵗʰ', '50ᵗʰ'), trans = 'reverse'
  ) +
  scale_color_manual(values = highlight_colors) +
  coord_cartesian(clip = 'off') +
  transition_reveal(year, keep_last = TRUE) +
  labs(
    title = 'The most popular national park_visits',
    subtitle = 'National park_visits ranked by number of visitors in a given year',
    x = '',
    y = 'Rank',
    caption = '#tidytuesday 38|2019  •  © 2019 spren9er'
  ) +
  theme(
    plot.background = element_rect(fill = '#f0f0f0'),
    plot.margin = margin(t = 40, r = 155, b = 20, l = 20),
    plot.title = element_text(
      margin = margin(b = 8), size = 38, hjust = -0.17, face = 'bold',
      color = '#333333'
    ),
    plot.subtitle = element_text(
      margin = margin(t = 6, b = 5), size = 29, hjust = -0.72,
      face = 'plain', color = '#333333'
    ),
    plot.caption = element_text(
      color = '#999999', size = 13, margin = margin(t = 10), hjust = 0.5,
      face = 'plain', family = 'Decima Mono Pro'
    ),
    panel.background = element_rect(fill = '#f0f0f0'),
    panel.grid.major = element_line(size = 0.5, color = '#d3d3d3'),
    panel.border = element_blank(),
    axis.text.x = element_text(
      family = 'Decima Mono Pro', color = '#999999', face = 'plain', size = 20,
      margin = margin(t = 6)
    ),
    axis.text.y = element_text(
      family = 'Decima Mono Pro', color = '#999999', face = 'plain', size = 20,
      margin = margin(r = 6)
    ),
    axis.title.y = element_text(color = '#333333', face = 'bold', size = 16)
  )

animate(animation, width = 1000, height = 1000, end_pause = 30)

anim_save('images/tidytuesday_201938_national_park_visits.gif')
```

## Year over Year Percent Change

source: https://github.com/gkaramanis/tidytuesday/blob/master/week-38/national-parks-lines.R

```{r year-over-year-change-in-total-visits-by-park}
pv_ch <- park_visits %>% 
  distinct(year, unit_name, unit_type, visitors) %>% 
  filter(unit_type == "National Park" & year != "Total") %>%
  mutate(year = as.numeric(year)) %>% 
  group_by(unit_name) %>% 
  arrange(year, .by_group = TRUE) %>%
  mutate(yoy = visitors - lag(visitors)) %>% 
  filter(unit_name != "Denali National Preserve")
  
 
pv_ch %>%
  ggplot() +
  geom_line(aes(x = year, y = yoy, group = unit_name), size = 0.2) +
  scale_x_continuous(breaks = seq(1910, 2010, 10), expand = expand_scale(add = c(5, 1))) +
  scale_y_continuous(breaks = c(-3000000, 0, 1000000), labels = paste0(c(-3000000, 0, 1000000)/1000000, "M")) +
  facet_wrap(vars(unit_name), ncol = 3) +
  labs(
    title = "National Park Visits 1904–2016",
    subtitle = "Year-over-year change in total visits by park",
    caption = "Source: dataisplural/data.world | Graphic: Georgios Karamanis"
  ) +
  theme_void(base_family = "IBM Plex Sans") +
  theme(
    legend.position = "none",
    legend.title = element_text(size = 20, color = "grey20"),
    legend.margin = margin(0, 0, 20, 0),
    plot.background = element_rect(fill = "grey80", color = NA),
    strip.background = element_rect(fill = "grey80", color = NA),
    strip.text = element_text(family = "IBM Plex Sans Bold", color = "grey30",
                               hjust = 1, vjust = 1),
    plot.title = element_text(size = 28, color = "grey20", family = "IBM Plex Sans Medium"),
    plot.subtitle = element_text(size = 20, color = "grey20", margin = margin(5, 0, 30, 0)),
    plot.caption = element_text(size = 8, color = "grey30", margin = margin(20, 0, 0, 0)),
    axis.text.x = element_text(family = "IBM Plex Mono", size = 7, color = "grey40"),
    axis.text.y = element_text(family = "IBM Plex Mono", size = 7, color = "grey40"),
    panel.grid.major.x = element_line(color = "grey75"),
    panel.grid.major.y = element_line(color = "grey75"),
    plot.margin = margin(20, 20, 20, 20)
  ) +

# save image
  ggsave(
  here::here("images", paste0("national-parks", format(Sys.time(), "%Y%m%d_%H%M%S"), ".png")),
  width = 18, height = 14, dpi = 320
)
```

## Mean National Park visits by year
source: https://github.com/gkaramanis/tidytuesday/blob/master/week-38/national-parks-mean-pct.R

```{r percentage-of-all-time-average-number-of-visits-by-year}
pv_ch <- park_visits %>% 
  distinct(year, unit_name, unit_type, visitors) %>% 
  filter(unit_type == "National Park" & year != "Total") %>%
  mutate(year = as.numeric(year)) %>% 
  group_by(unit_name) %>% 
  filter(unit_name != "Denali National Preserve") %>% 
  mutate(
    mean_visitors = mean(visitors),
    mean_pct = log10(visitors/mean_visitors)
    ) %>% 
  add_tally()


ggplot(pv_ch) +
  geom_segment(aes(x = year, xend = year,
                   y = 0, yend = 0.5, color = cut_number(mean_pct, 10)), size = 1.4) +
  scale_x_continuous(breaks = seq(1910, 2010, 20), expand = expand_scale(add = c(5, 1))) +
  scale_y_continuous(expand = c(0.05, 0.25)) +
  facet_wrap(vars(fct_reorder(unit_name, -n)), ncol = 3) +
  #scale_color_viridis(option = "plasma", discrete = TRUE) +
  labs(
    title = "National Park Visits, 1904–2016",
    subtitle = "Percentage of all-time average number of visits, by year",
    caption = "Source: dataisplural/data.world | Graphic: Georgios Karamanis"
  ) +
  guides(color = guide_legend(
    # title.position = "top",
    label.position = "bottom",
    nrow = 1,
    title = NULL,
    barwidth = 20,
    barheight = 0.5
    )) +
  theme_void(base_family = "IBM Plex Sans") +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 20, color = "grey20"),
    legend.margin = margin(0, 0, 20, 0),
    plot.background = element_rect(fill = "grey80", color = NA),
    strip.background = element_rect(fill = "grey80", color = NA),
    strip.text = element_text(family = "IBM Plex Sans Bold", color = "grey30",
                               hjust = 1, vjust = 1),
    plot.title = element_text(size = 28, color = "grey20", family = "IBM Plex Sans Medium"),
    plot.subtitle = element_text(size = 20, color = "grey20"),
    plot.caption = element_text(size = 8, color = "grey30", margin = margin(20, 0, 0, 0)),
    axis.text.x = element_text(family = "IBM Plex Mono", size = 7, color = "grey40"),
    panel.grid.major.x = element_line(color = "grey75"),
    plot.margin = margin(20, 20, 20, 20)
  ) +

# save image
  ggsave(
  here::here("images", paste0("national-parks", format(Sys.time(), "%Y%m%d_%H%M%S"), ".png")),
  width = 18, height = 14, dpi = 320
)
```

## National Parks visits
Source: https://github.com/gkaramanis/tidytuesday/blob/master/week-38/national-parks.R

```{r percentage-of-all-time-average-number-of-visits-by-year-2}
pv_ch <- park_visits %>% 
  distinct(year, unit_name, unit_type, visitors) %>% 
  filter(unit_type == "National Park" & year != "Total") %>%
  mutate(year = as.numeric(year)) %>% 
  group_by(unit_name) %>% 
  filter(unit_name != "Denali National Preserve") %>% 
  mutate(
    mean_visitors = mean(visitors),
    mean_pct = log10(visitors/mean_visitors),
    mean_pct_cut = cut_interval(mean_pct, 10)
  ) %>% 
  add_tally()


ggplot(pv_ch) +
  geom_segment(aes(x = year, xend = year,
                   y = 0, yend = 0.5, color = mean_pct_cut), size = 1.4) +
  scale_x_continuous(breaks = seq(1910, 2010, 20), expand = expand_scale(add = c(5, 1))) +
  scale_y_continuous(expand = c(0.05, 0.25)) +
  facet_wrap(vars(fct_reorder(unit_name, -n)), ncol = 3) +
  scale_colour_viridis_d(option = "plasma") +
  labs(
    title = "National Park Visits, 1904–2016",
    subtitle = "Percentage of all-time average number of visits, by year",
    caption = "Source: dataisplural/data.world | Graphic: Georgios Karamanis"
  ) +
  guides(color = guide_colorbar(
    title.position = "top",
    label.position = "top",
    title = NULL,
    barwidth = 20,
    barheight = 0.5
  )) +
  theme_void(base_family = "IBM Plex Sans") +
  theme(
    legend.position = "none",
    legend.title = element_text(size = 20, color = "grey20"),
    legend.margin = margin(0, 0, 20, 0),
    plot.background = element_rect(fill = "grey80", color = NA),
    strip.background = element_rect(fill = "grey80", color = NA),
    strip.text = element_text(family = "IBM Plex Sans Bold", color = "grey30",
                              hjust = 1, vjust = 1),
    plot.title = element_text(size = 28, color = "grey20", family = "IBM Plex Sans Medium"),
    plot.subtitle = element_text(size = 20, color = "grey20"),
    plot.caption = element_text(size = 8, color = "grey30", margin = margin(20, 0, 0, 0)),
    axis.text.x = element_text(family = "IBM Plex Mono", size = 7, color = "grey40"),
    panel.grid.major.x = element_line(color = "grey75"),
    plot.margin = margin(20, 20, 20, 20)
  ) +
  
  # save image
  ggsave(
    here::here("images", paste0("national-parks", format(Sys.time(), "%Y%m%d_%H%M%S"), ".png")),
    width = 18, height = 14, dpi = 320
  )
```

## Add park fees
Source: https://github.com/jkaupp/tidytuesdays/blob/master/2019/week38/R/analysis.R

```{r get-park-fees}
# Get park fees
fees_page <- "https://www.nps.gov/aboutus/entrance-fee-prices.htm"

parks <- read_html(fees_page) %>% 
  html_nodes("h3") %>% 
  html_text() %>% 
  .[-1:-2]

park_fees <- read_html(fees_page) %>% 
  html_nodes(".table-wrapper > table") %>% 
  html_table() %>% 
  map(~set_names(.x, c("date", "park_specific_annual_pass", "per_vehicle", "per_person", "per_motorcycle"))) %>% 
  map2(parks, ~mutate(.x, park = .y)) %>% 
  bind_rows() %>% 
  filter(date == "Current") %>% 
  rename(park_name = park) %>% 
  mutate(park_name = stringi::stri_trans_general(park_name, id = "Latin-ASCII"),
         park_name = str_replace(park_name, "Hawai'i", "Hawaii"))
```

### Read in updated data for annual summary for national parks

```{r read-updated-data, message=F, warning=F}
summary_report <- read_csv(here("data/annual_summary_report.csv")) %>% clean_names()
```

## Plot estimated national park revenue from fees for 2018

```{r estimated-national-park-revenue-from-fees-for-2018}
park_revenue_data <- summary_report %>% 
  filter(year == 2018) %>% 
  mutate(visitors = recreation_visitors + non_recreation_visitors) %>% 
  select(year, park_name, visitors) %>% 
  mutate(park_name = str_remove(park_name, "[A-Z]{2,}"),
         park_name = str_remove(park_name, "& PRES"),
         park_name = trimws(park_name)) %>% 
  regex_left_join(park_fees, ., ignore_case = TRUE) %>% 
  distinct(year, park_name.x, .keep_all = TRUE) %>% 
  filter(str_detect(park_name.x, "Park"), !str_detect(park_name.x, "Great Falls")) %>% 
  mutate(revenue = visitors * parse_number(per_person)) %>% 
  rename(park_name = park_name.x) %>% 
  select(-park_name.y)


plot <- ggplot(park_revenue_data, aes(x = fct_reorder(park_name, revenue), y = revenue)) +
  geom_col(fill = "#5e81ac", size = 0.1) +
  coord_flip() +
  scale_y_continuous(labels = scales::dollar, expand = c(0.01,0)) +
  labs(title = "Estimated National Park Revenue from Fees for 2018",
       subtitle = str_wrap("Illustrated below is a bar chart of fee revenue from US National Parks in 2018.  Estimated Revenue calculated using per person admittance rates and total park visitors.", 95),
       caption = "Data: www.nps.gov | Graphic: @jakekaupp",
       x = NULL,
       y = NULL) +
  theme_jk(grid = "X") +
  theme(plot.background = element_rect(fill = "#2e3440"),
        text = element_text(color = "#eceff4"),
        panel.grid = element_line(color = "#e5e9f0"),
        axis.text.x = element_text(color = "#eceff4"),
        axis.text.y = element_text(color = "#eceff4"))

ggsave(here("images", "tw_38plot.png"), plot, width = 10, height = 8, device = agg_png())
```

## Getting location from park names via ggmap
Using the park name, or any location name, and the Geocoding API from Google, we got the longitude and latitude for each parks locations.

```{r getting-location-from-park-names-via-ggmap, message=F}
# library(ggmap)
# unique_locations <- summary_report %>% distinct(park_name)
# locations_df <- mutate_geocode(unique_locations, park_name)
# write_csv(locations_df, 'data/park_locations_from_google_api.csv')
```

## Getting data for geocoded plot

```{r adding-location-data-for-plot, message=F}
trial <- summary_report %>% 
  mutate(visitors = recreation_visitors + non_recreation_visitors) %>% 
  select(year, park_name, visitors)

park_locations <- read_csv('data/park_locations_from_google_api.csv')
colnames(park_locations)[colnames(park_locations)=="lon"] <- "long"

plot_data <- trial %>% 
  regex_left_join(park_locations, ., ignore_case = TRUE)

plot_data <- plot_data %>% filter(long > -128, lat < 51, lat > 24)
plot_data <- plot_data[complete.cases(plot_data), ] %>% group_by(year, park_name.x)

split_by_year <- plot_data %>% split(plot_data$year)

park_names_over_the_years <- split_by_year$`1979`$park_name.x
for (i in split_by_year) {
  park_names_over_the_years <- intersect(park_names_over_the_years, i$park_name.x)
}

plot_data <- plot_data %>% filter(park_name.x %in% park_names_over_the_years)
plot_data$year <- as.integer(plot_data$year)
plot_data$visitors <- scale(plot_data$visitors)
```

## Building plot for park visits through the years on the US map

```{r us-park-visits-through-the-years, message=F}
library(maps)
us_states <- map_data("usa")

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

g <- ggplot(data = us_states) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill=NA, color= 'black') + 
  coord_fixed(1.3) +
  geom_point(data = plot_data, aes(x = long, y = lat, color = visitors, size = visitors)) +
  guides(size=FALSE) +
  scale_colour_gradient(low = "blue", high="red") +
  ditch_the_axes

g + transition_time(year) +
  labs(title = "Year: {frame_time}, Park visits, scaled")
g

anim_save("images/park_visits_by_year.gif")
```

